---
title: 'Peer-graded Assignment: R Markdown and Leaflet'
output:
  html_document:
    keep_md: yes
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

This assignment is to showcase one of many use cases of the `leaflet` interactive mapping visualization package. We will discuss 5 of the big cities in China to implement population visualisation in this case. 

## Required Libraries

The following packages are required to reproduce.

```{r 'library', eval=T, warning=FALSE, message=F}
require(ggmap); require(leaflet); require(magrittr); require(data.table)
```

## Data Preparation

The total area (km<sup>2</sup>) and population density (/km<sup>2</sup>) data are acquired from [wikipedia](https://en.wikipedia.org/wiki/List_of_cities_proper_by_population). 
First we load the data into a dataframe.

```{r 'data_preparation'}
City <- c('Shanghai', 'Beijing', 'Guangzhou', 'Nanjing', 'Chongqing')

Total.area <- c(6340.5, 16410.54, 7434.4, 4713.85, 5473)

Pop.dens <- c(3826, 1311, 1759, 1737, 1496)
```

Next we need to obtain the coordinate of the cities. This is being done by leveraging on `ggmap` package. However, the google API will be triggered everytime the code is being executed. Therefore, we will just manually copy the result from the API and save it in R.

```{r 'get_coord'}
lon <- c(121.47583008, 116.41113281, 113.27041626, 118.80065918, 106.92443848)

lat <- c(31.21749936, 39.8928799, 23.12394255, 32.05464469, 29.41567547)
```

Now we can create a dataframe that contains all the information that we need to visualize it in `leaflet`.

```{r 'create_dataframe'}
mydata <- data.table(city = City, total.area = Total.area, pop.dens = Pop.dens, lon = lon, lat = lat)

## Transform the population density value to the range of 0 and 1 for color mapping
mydata[, color:= (pop.dens - min(pop.dens))/diff(range(pop.dens))]

## Create the pop up description and urbanisation rate
mydata[, popup:= paste0(city, '<br>Population Density: ', pop.dens, '/km2', '<br>Total Area ', total.area, 'km2')]
```

## Visualization in Leaflet

Visualize the population data in leaflet.

```{r 'visualise'}

## Define a color function to map the population density to a color

f.color <- colorRamp(c(c('green', 'red')))

leaflet(mydata) %>% addTiles() %>% addCircles(weight = 1, radius = 30000, color = f.color(mydata$color) %>% rgb(maxColorValue = 256)) %>% addMarkers(popup = mydata$popup)
```