---
title: 'Statistical Inference: Peer Assessment Part 2'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Requirements

+ Load the ToothGrowth data and perform some basic exploratory data analyses
+ Provide a basic summary of the data.
+ Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
+ State your conclusions and the assumptions needed for your conclusions.

## Including Libraries
```{r}
library(ggplot2)
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
```
## Question 1
Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
plot <- ggplot(ToothGrowth, 
               aes(x = factor(dose),
               y = len,
               fill = factor(dose)))
               
plot +
geom_boxplot(notch = F) + 
facet_grid(.~supp) +
scale_x_discrete("Dosage (mg)") +   
scale_y_continuous("Length of Teeth") +  
ggtitle("Exploratory Data Analyses")
```

## Question 2
Provide a basic summary of the data.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
```
```{r}
table(ToothGrowth$supp, ToothGrowth$dose)
```

## Question 3
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r}
supp.t1 <- t.test(len~supp, paired=F, var.equal=T, data=ToothGrowth)
supp.t2 <- t.test(len~supp, paired=F, var.equal=F, data=ToothGrowth)
supp.result <- data.frame("p-value" = c(supp.t1$p.value, supp.t2$p.value),
                          "ConfLow" = c(supp.t1$conf[1],supp.t2$conf[1]),
                          "Conf-High" = c(supp.t1$conf[2],supp.t2$conf[2]),
                          row.names = c("Equal Var","Unequal Var"))
supp.result
```

## Question 4
State your conclusions and the assumptions needed for your conclusions.

Based on the analysis above, our conclusions are as follows: 

- The 2mg dose has larger impact on tooth growth than 1mg and 0.5mg, while 1mg dose has more impact than 0.5mg dose. So there is a different in the growth of the tooth while the doses are larger.
- Orange juice and vitamin C have obvious different impact on tooth growth.